---
title: 'Prophet Mbonye Interview with Serwanja '
author: ''
date: '2020-04-28'
slug: prophet-mbonye-interview-with-serwanja
categories:
  - Analysis
tags:
  - R
  - tidytexts
lastmod: '2020-04-28T14:37:28+03:00'
layout: post
type: post
highlight: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

**Elvis Mbonye Claims to have prophesied Corona Virus?? But that not the Point of this Blog.**

Am just interested in what happened during the interview, what words were said, were they Positive or Negative. Did the Interview Spark any emotions like joy, anger, sadness, and so forth?? 
That's what am going to be analysing here.

Note : This isn't complete until the all the episodes of the interview are out.

```{r}
library(tidyverse)
library(tidytext)
library(wordcloud)
```

### How did i get the data??
After a little research online, i realised i could scrape the auto generated subtitles by youtube and download them as a txt file then read it in excel. its not perfect but it works and its better than transcribing the audio which is another option.

```{r}
data <- readr::read_lines(file.choose()) %>% as_tibble()
```

### unnesting words to a single row each and remove stop words 
```{r}
data(stopwords)

text_data <- data %>% 
        unnest_tokens(word, value) %>% 
        anti_join(stop_words)
        
```

### What is the most common word during the interview??
```{r}
text_data %>% 
        count(word, sort = TRUE) %>% 
        head(20) %>% 
        mutate(word = fct_reorder(word, n)) %>% 
        ggplot(aes(word, n)) + 
        geom_col() + 
        coord_flip() +
        labs(y="# of words")

text_data %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))
```
