---
title: NA == NA = TRUE? Wrong
author: ''
date: '2020-04-08'
slug: na-na-na
categories:
  - eli5
tags:
  - NA
  - Nan
  - R
lastmod: '2020-04-08T11:18:33+03:00'
layout: post
type: post
highlight: no
---

### Why is NA == NA = NA?
Most Objects in R compared to themselves return **TRUE** but no **NA**.
It's actually i had never thought about but thinking about it makes some alot of sense, Comparing two Unknowns shouldn't should be return TRUE.

Let's Look at some of the Most Common examples that return *TRUE*

```{r}
10 == 10
```
```{r}
-10 == -10
```
```{r}
FALSE == FALSE
```

```{r}
"NA" == "NA"
```
This is a string **"NA"** and not the **NA** we are talking about and will return a TRUE just like any other similar string comparison

## So Why

Think about it this way, if i say i have **$NA** in my bank account and you also have **$NA** in your bank account. I can't confidently say we have an equal in our bank accounts because we i don't know how much you got and so do you. 

So a comparison of Our Bank account balances is surely Unknown. Hence **NA**
```{r}

NA == NA

```

[Read More about NA's in R4DS](https://r4ds.had.co.nz/)





















